Poster Analysis

### 2024-10-24 Manuscript Analysis/ASHA Poster Analyses.Rmd
rendered r markdown file

---
title: "Poster Analysis"
author: "Micah Hirsch"
date: "2024-10-21"
output: html_document
---

@@

This is the best fitting model for the pupil dilation data. In regards to the cognitive data, here are the key take-aways:

@@

ple |>
  dplyr::select(subject, trial, speaker, code, effort_rating, rep_acc, ends_with("_c")) |>
  dplyr::group_by(subject, speaker, rep_acc, list_sort_c, flanker_c, card_sort_c, pattern_c) |>
  dplyr::summarize(effort = mean(effort_rating)) |>
  dplyr::rename('Working Memory' = list_sort_c,
                'Inhibitory Control' = flanker_c,
                'Cognitive Flexibility' = card_sort_c,
                'Processing Speed' = pattern_c) |>
  tidyr::pivot_longer(cols = "Working Memory":"Processing Speed",
                     names_to = "cog_measure",
                     values_to = "score") |>
  ungroup() |>
  ggplot() +
  aes(x = score,
      y = effort,
      color = speaker,
      shape = speaker) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm") +
  facet_grid(cog_measure ~ rep_acc) +
  labs(x = "Score", y = "Perceived Listening Effort Rating") +
  theme_classic() +
  scale_color_manual(values = my_pal) +
  scale_shape_manual(values = c(1, 2)) +
  theme(aspect.ratio = 1,
        panel.spacing.x = unit(.8, "lines"),
        legend.position = "bottom")

@@

m0 <- glmmTMB(effort_rating ~ 1 + (1|subject) + (1|code), data = ple)
sjPlot::tab_model(m0, pred.labels = "Intercept",
                  dv.labels = "Perceived Listening Effort")

performance::icc(m0)

@@

>>>>

---
title: "Poster Analysis"
author: "Micah Hirsch"
date: "2024-10-21"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

@@

This is the best fitting model for the pupil dilation data. Model assumptions were also visually checked using the performance package. In regards to the cognitive data, here are the key take-aways:

@@

ple |>
  dplyr::select(subject, trial, speaker, code, effort_rating, rep_acc, ends_with("_c")) |>
  dplyr::group_by(subject, speaker, rep_acc, list_sort_c, flanker_c, card_sort_c, pattern_c) |>
  dplyr::summarize(effort = mean(effort_rating)) |>
  dplyr::rename('Working Memory' = list_sort_c,
                'Inhibitory Control' = flanker_c,
                'Cognitive Flexibility' = card_sort_c,
                'Processing Speed' = pattern_c) |>
  tidyr::pivot_longer(cols = "Working Memory":"Processing Speed",
                     names_to = "cog_measure",
                     values_to = "score") |>
  ungroup() |>
  ggplot() +
  aes(x = score,
      y = effort,
      color = speaker,
      shape = speaker) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm") +
  facet_grid(cog_measure ~ rep_acc) +
  labs(x = "Score", y = "Perceived Listening Effort Rating") +
  theme_classic() +
  scale_color_manual(values = my_pal, labels = c("Neurotypical", "ALS")) +
  scale_shape_manual(values = c(1, 2), labels = c("Neurotypical", "ALS")) +
  theme(aspect.ratio = 1,
        panel.spacing.x = unit(.8, "lines"),
        legend.position = "bottom")

@@

m0 <- glmmTMB(effort_rating ~ 1 + (1|subject) + (1|code), data = ple)
sjPlot::tab_model(m0, 
                  pred.labels = "Intercept",
                  dv.labels = "Perceived Listening Effort")
performance::icc(m0)
performance::check_model(m0)

@@

m2 <- glmmTMB(effort_rating ~ trial_c + speaker*rep_acc + (trial_c|subject) + (trial_c|code), data = ple)
sjPlot::tab_model(m2,
                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]",
                                  "Repetition Accuracy [Inaccurate]", "Speaker * Repetition Accuracy"),
                  dv.labels = "Perceived Listening Effort")

performance::test_performance(m1, m2)

m3 <- glmmTMB(effort_rating ~ trial_c + speaker*rep_acc + scale(list_sort_c, scale = F) +(trial_c|subject) + (trial_c|code), data = ple)
sjPlot::tab_model(m3,
                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]",
                                  "Repetition Accuracy [Inaccurate]", "Working Memory",
                                  "Speaker * Repetition Accuracy"),
                  dv.labels = "Perceived Listening Effort")

performance::test_performance(m2, m3)

m4 <- glmmTMB(effort_rating ~ trial_c + speaker*rep_acc + scale(flanker_c, scale = F) +(trial_c|subject) + (trial_c|code), data = ple)
sjPlot::tab_model(m4,
                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]",
                                  "Repetition Accuracy [Inaccurate]", "Inhibitory Control",
                                  "Speaker * Repetition Accuracy"),
                  dv.labels = "Perceived Listening Effort")

performance::test_performance(m2, m4)

m5 <- glmmTMB(effort_rating ~ trial_c + speaker*rep_acc + scale(card_sort_c, scale = F) +(trial_c|subject) + (trial_c|code), data = ple)
sjPlot::tab_model(m5,
                  pred.labels = c("Intercept", "Trial Number", "Spekaer [ALS]",
                                  "Repetition Accuracy [Inaccurate]", "Cognitive Flexibility",
                                  "Speaker * Repetition Accuracy"),
                  dv.labels = "Perceived Listening Effort")

performance::test_performance(m2, m5)

m6 <- glmmTMB(effort_rating ~ trial_c + speaker*rep_acc + scale(pattern_c, scale = F) +(trial_c|subject) + (trial_c|code), data = ple)
sjPlot::tab_model(m6,
                  pred.labels = c("Intercept", "Trial Number", "Spekaer [ALS]",
                                  "Repetition Accuracy [Inaccurate]", "Processing Speed",
                                  "Speaker * Repetition Accuracy"),
                  dv.labels = "Perceived Listening Effort")

performance::test_performance(m2, m6)

m7 <- glmmTMB(effort_rating ~ trial_c + speaker*rep_acc + 
                speaker*scale(list_sort_c, scale = F) + (trial_c|subject) + (trial_c|code), data = ple)
sjPlot::tab_model(m7,
                  pred.labels = c("Intercept", "Trial Number", "Spekaer [ALS]",
                                  "Repetition Accuracy [Inaccurate]", "Working Memory",
                                  "Speaker * Repetition Accuracy",
                                  "Working Memory * Speaker [ALS]"),
                  dv.labels = "Perceived Listening Effort")

performance::test_performance(m2, m7)

m8 <- glmmTMB(effort_rating ~ trial_c + speaker*rep_acc + 
                speaker*scale(list_sort_c, scale = F) + speaker*scale(flanker_c, scale = F) + 
                (trial_c|subject) + (trial_c|code), data = ple)
sjPlot::tab_model(m8,
                  pred.labels = c("Intercept", "Trial Number", "Spekaer [ALS]",
                                  "Repetition Accuracy [Inaccurate]", "Working Memory", "Inhibitory Control",
                                  "Speaker * Repetition Accuracy",
                                  "Working Memory * Speaker [ALS]", "Inhibitory Control * Speaker [ALS]"),
                  dv.labels = "Perceived Listening Effort")

performance::test_performance(m7, m8)

```
### Model 9

m9 <- glmmTMB(effort_rating ~ trial_c + speaker*rep_acc + 
                speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) + 
                (trial_c|subject) + (trial_c|code), data = ple)
sjPlot::tab_model(m9,
                  pred.labels = c("Intercept", "Trial Number", "Spekaer [ALS]",
                                  "Repetition Accuracy [Inaccurate]", "Working Memory", "Cognitive Flexibility",
                                  "Speaker * Repetition Accuracy",
                                  "Working Memory * Speaker [ALS]", "Cognitive Flexibility * Speaker [ALS]"),
                  dv.labels = "Perceived Listening Effort")

performance::test_performance(m7, m9)

m10 <- glmmTMB(effort_rating ~ trial_c + speaker*rep_acc + 
                speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) + 
                 speaker*scale(pattern_c, scale = F) +
                (trial_c|subject) + (trial_c|code), data = ple)
sjPlot::tab_model(m10, 
                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]",
                                  "Repetition Accuracy [Inaccurate]", "Working Memory", "Cognitive Flexibility",
                                  "Processing Speed",
                                  "Speaker * Repetition Accuracy",
                                  "Working Memory * Speaker [ALS]", "Cognitive Flexibility * Speaker [ALS]"),
                  dv.labels = "Perceived Listening Effort")

performance::test_performance(m9, m10)

```
### Model 11

m11 <- glmmTMB(effort_rating ~ trial_c + speaker*rep_acc + 
                 speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) +
                 rep_acc*scale(list_sort_c, scale = F) +
                (trial_c|subject) + (trial_c|code), data = ple)
sjPlot::tab_model(m11, 
                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]",
                                  "Repetition Accuracy [Inaccurate]", "Working Memory", "Cognitive Flexibility",
                                  "Speaker * Repetition Accuracy",
                                  "Working Memory * Speaker [ALS]", "Cognitive Flexibility * Speaker [ALS]", 
                                  "Working Memory * Repetition Accuracy [Inaccurate]"),
                  dv.labels = "Perceived Listening Effort")

performance::test_performance(m9, m11)

m12 <- glmmTMB(effort_rating ~ trial_c + speaker*rep_acc + 
                 speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) +
                 rep_acc*scale(list_sort_c, scale = F) + rep_acc*scale(flanker_c, scale = F) +
                (trial_c|subject) + (trial_c|code), data = ple)
sjPlot::tab_model(m12, 
                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]",
                                  "Repetition Accuracy [Inaccurate]", "Working Memory", "Cognitive Flexibility",
                                  "Inhibitory Control", "Speaker * Repetition Accuracy",
                                  "Working Memory * Speaker [ALS]", "Cognitive Flexibility * Speaker [ALS]", 
                                  "Working Memory * Repetition Accuracy [Inaccurate]",
                                  "Inhibitory Control * Repetition Accuracy [Inaccurate]"),
                  dv.labels = "Perceived Listening Effort")

performance::test_performance(m11, m12)

m13 <- glmmTMB(effort_rating ~ trial_c + speaker*rep_acc + 
                 speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) +
                 rep_acc*scale(list_sort_c, scale = F) + rep_acc*scale(card_sort_c, scale = F) +
                (trial_c|subject) + (trial_c|code), data = ple)
sjPlot::tab_model(m13, 
                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]",
                                  "Repetition Accuracy [Inaccurate]", "Working Memory", "Cognitive Flexibility",
                                  "Speaker * Repetition Accuracy",
                                  "Working Memory * Speaker [ALS]", "Cognitive Flexibility * Speaker [ALS]", 
                                  "Working Memory * Repetition Accuracy [Inaccurate]",
                                  "Cognitive Flexibility * Repetition Accuracy [Inaccurate]"),
                  dv.labels = "Perceived Listening Effort")

performance::test_performance(m11, m13)

```
### Model 14

m14 <- glmmTMB(effort_rating ~ trial_c + speaker*rep_acc + 
                 speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) +
                 rep_acc*scale(list_sort_c, scale = F) + rep_acc*scale(card_sort_c, scale = F) +
                 rep_acc*scale(pattern_c, scale = F) +
                (trial_c|subject) + (trial_c|code), data = ple)
sjPlot::tab_model(m14, 
                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]",
                                  "Repetition Accuracy [Inaccurate]", "Working Memory", "Cognitive Flexibility",
                                  "Processing Speed", "Speaker * Repetition Accuracy",
                                  "Working Memory * Speaker [ALS]", "Cognitive Flexibility * Speaker [ALS]", 
                                  "Working Memory * Repetition Accuracy [Inaccurate]",
                                  "Cognitive Flexibility * Repetition Accuracy [Inaccurate]",
                                  "Processing Speed * Repetition Accuracy [Inaccurate]"),
                  dv.labels = "Perceived Listening Effort")

performance::test_performance(m13, m14)

m15 <- glmmTMB(effort_rating ~ trial_c + speaker*rep_acc + 
                 speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) +
                 rep_acc*scale(list_sort_c, scale = F) + rep_acc*scale(card_sort_c, scale = F) +
                 speaker*rep_acc*scale(list_sort_c, scale = F) +
                (trial_c|subject) + (trial_c|code), data = ple)
sjPlot::tab_model(m15, 
                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]",
                                  "Repetition Accuracy [Inaccurate]", "Working Memory", "Cognitive Flexibility", 
                                  "Speaker * Repetition Accuracy",
                                  "Working Memory * Speaker [ALS]", "Cognitive Flexibility * Speaker [ALS]", 
                                  "Working Memory * Repetition Accuracy [Inaccurate]",
                                  "Cognitive Flexibility * Repetition Accuracy [Inaccurate]",
                                  "Working Memory * Speaker * Accuracy"),
                  dv.labels = "Perceived Listening Effort")

performance::test_performance(m13, m15)

m16 <- glmmTMB(effort_rating ~ trial_c + speaker*rep_acc + 
                 speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) +
                 rep_acc*scale(list_sort_c, scale = F) + rep_acc*scale(card_sort_c, scale = F) +
                 speaker*rep_acc*scale(flanker_c, scale = F) +
                (trial_c|subject) + (trial_c|code), data = ple)
sjPlot::tab_model(m16, 
                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]",
                                  "Repetition Accuracy [Inaccurate]", "Working Memory", "Cognitive Flexibility", 
                                  "Inhibitory Control", "Speaker * Repetition Accuracy",
                                  "Working Memory * Speaker [ALS]", "Cognitive Flexibility * Speaker [ALS]", 
                                  "Working Memory * Repetition Accuracy [Inaccurate]",
                                  "Cognitive Flexibility * Repetition Accuracy [Inaccurate]",
                                  "Inhibitory Control * Speaker [ALS]", 
                                  "Inhibitory Control * Repetition Accuracy [Inaccurate]",
                                  "Inhibitory Control * Speaker * Accuracy"),
                  dv.labels = "Perceived Listening Effort")

performance::test_performance(m13, m16)

m17 <- glmmTMB(effort_rating ~ trial_c + speaker*rep_acc + 
                 speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) +
                 rep_acc*scale(list_sort_c, scale = F) + rep_acc*scale(card_sort_c, scale = F) +
                 speaker*rep_acc*scale(card_sort_c, scale = F) +
                (trial_c|subject) + (trial_c|code), data = ple)
sjPlot::tab_model(m17, 
                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]",
                                  "Repetition Accuracy [Inaccurate]", "Working Memory", "Cognitive Flexibility", 
                                  "Speaker * Repetition Accuracy",
                                  "Working Memory * Speaker [ALS]", "Cognitive Flexibility * Speaker [ALS]", 
                                  "Working Memory * Repetition Accuracy [Inaccurate]",
                                  "Cognitive Flexibility * Repetition Accuracy [Inaccurate]",
                                  "Cognitive Flexibility * Speaker * Accuracy"),
                  dv.labels = "Perceived Listening Effort")

performance::test_performance(m13, m17)

```{r,warning = F, message = F}

m18 <- glmmTMB(effort_rating ~ trial_c + speaker*rep_acc + 
                 speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) +
                 rep_acc*scale(list_sort_c, scale = F) + rep_acc*scale(card_sort_c, scale = F) +
                 speaker*rep_acc*scale(pattern_c, scale = F) +
                (trial_c|subject) + (trial_c|code), data = ple)
sjPlot::tab_model(m18, 
                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]",
                                  "Repetition Accuracy [Inaccurate]", "Working Memory", "Cognitive Flexibility", 
                                  "Processing Speed", "Speaker * Repetition Accuracy",
                                  "Working Memory * Speaker [ALS]", "Cognitive Flexibility * Speaker [ALS]", 
                                  "Working Memory * Repetition Accuracy [Inaccurate]",
                                  "Cognitive Flexibility * Repetition Accuracy [Inaccurate]",
                                  "Processing Speed * Speaker [ALS]",
                                  "Processing Speed * Repetition Accuracy [Inaccurate]",
                                  "Processing Speed * Speaker * Accuracy"),
                  dv.labels = "Perceived Listening Effort")

performance::test_performance(m13, m18)

Based on the model-building steps above, model 13 is the best fitting model for the perceived listening effort ratings. Since the interaction between speaker and repetition accuracy was not significant during any model building steps, that was removed from the model. Removal of this interaction did not significantly impact model performance. Model assumptions were also visually checked. Most assumptions appear to be met.

The final model take-aways, in reference to the cognitve measures of interest, are:
- The interaction between working memory and speaker: On average, when listening to the speaker with ALS, listeners with higher working memory scores rated effort higher. 
- The interaction between cognitive flexibility and speaker: On average, when listening to the speaker with ALS, listeners with higher cognitive flexibility scores rated effort lower.
- The interaction between working memory and repetition accuracy: On average, listeners with higher working memory scores rated effort phrases they did not accurately recognized lower.
- The interaction between cognitive flexibility and repetition accuracy: On average, listeners with higher cognitive flexibility scores rated effort higher for phrases they did not accurately recognize.
- Teh three way interactions between speaker, repetition accuracy, and each of the cognitve measures was tested. None of these interactions were significant.

m13b <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + 
                 speaker*scale(list_sort_c, scale = F) + speaker*scale(card_sort_c, scale = F) +
                 rep_acc*scale(list_sort_c, scale = F) + rep_acc*scale(card_sort_c, scale = F) +
                (trial_c|subject) + (trial_c|code), data = ple)
sjPlot::tab_model(m13b, 
                  pred.labels = c("Intercept", "Trial Number", "Speaker [ALS]",
                                  "Repetition Accuracy [Inaccurate]", "Working Memory", "Cognitive Flexibility",
                                  "Working Memory * Speaker [ALS]", "Cognitive Flexibility * Speaker [ALS]", 
                                  "Working Memory * Repetition Accuracy [Inaccurate]",
                                  "Cognitive Flexibility * Repetition Accuracy [Inaccurate]"),
                  dv.labels = "Perceived Listening Effort",
                  show.std = T,
                  file = "Tables/ple_model.html")

performance::test_performance(m13, m13b)
performance::check_model(m0)
